Out-of-Sample Comparisons of Overfit Models

### 2014-03-21 paper.tex
Revise Monte Carlo presentation
This commit edits the presentation of the Monte Carlo results,
hopefully making them seem more important and making them
easier to follow.

It also fixes several bugs in the graphs and converts the
presentation of the F-test results to a table.

Consider two sequences of length $P$ of prediction errors, the result
of forecasting the same variable with two different estimated models.
Both models are estimated with $R$ observations, collectively called
the {\em estimation window}, and are used to forecast an additional
$P$ observations, called the {\em test sample}.  There are $T$
observations in all, and $R+P=T$.  This paper introduces a new limit
theory for statistics constructed from these prediction errors
designed to approximate the behavior of the statistics when one of the
models is overfit.  In doing so, we provide a theoretical
justification for forecasters to use OOS instead of in-sample
comparisons: the DMW OOS test%
\footnote{In this paper, we will refer to the basic OOS $t$-test
  studied by \citet{DiM:95} and \citet{Wes:96} as the DMW test.} %
allows a forecaster to conduct
inference about the expected future accuracy of his or her models when
one or both is overfit.  We show analytically and through Monte Carlo
that standard full-sample test statistics can not test hypotheses
about this performance.

@@

We focus on linear regression models estimated with a fixed window
for simplicity, but our basic conclusions should be true for other
models and estimation strategies as well.  Under this
asymptotic theory, where the number of regressors $k$ increases with $T$
so that $\lim k/T$ is positive and less than one,
the OLS coefficient estimator is no longer
consistent or asymptotically normal \citep{Hub:73} and has positive
variance in the limit.  We show that, even so, the usual OOS average
is asymptotically normal and can consistently estimate the difference
between the models' generalization error, the expected loss in the
future conditional on the on the data available in period
$T$.%
\footnote{See, for example, \citet{HTF:08} for a discussion of
  generalization error.} %
Under these asymptotics, the generalization
error does not converge to the expected performance of the pseudotrue
models, so existing in-sample and OOS comparisons measure different
quantities and should be expected to give different results for
reasons beyond simple size and power comparisons.  Under our limit theory,
the model that is closer to the true DGP in population can forecast
worse.  In such a situation, a standard in-sample comparison would
correctly reject the null hypothesis that the benchmark is true, and
an OOS comparison would correctly fail to reject the null that the
benchmark is more accurate.%
\footnote{In a pair of papers similar to
  ours, \citet{ClM:12,ClM:12b} study in-sample and OOS tests that
  the larger model has nonzero coefficients that are too close to zero
  to expect the model to forecast more accurately.  Like this paper,
  they argue that the larger model can be true but less accurate.
  However, they focus on an aspect of the DGP that makes this
  phenomenon likely, while we focus on the coefficient estimates that
  produce less accurate forecasts.  Moreover, the implications of our
  asymptotic theories are different and their papers do not provide
  reasons to do OOS comparisons.} %
Also note that, in this situation, a model that performs well
in-sample can perform badly out-of-sample even if there are no
structural breaks or other forms of instability.  Researchers
have argued that a breakdown of in-sample forecasting ability
indicates a structural break (see \citealp{BoH:99}, and
\citealp{StW:03}, among others), but this breakdown can be caused by
overfit as well.

These theoretical results partially justify OOS comparisons when
researchers want to
choose a model for forecasting.  Although there has been little
emphasis on hypothesis testing in this setting, testing is usually
appropriate: there is usually a familiar benchmark model in place, and
the cost of incorrectly abandoning the benchmark for a less accurate
alternative model is higher than the cost of incorrectly failing to
switch to a more accurate alternative.  We show that the DMW test
lets the forecaster control the probability of the first error, just
as with conventional hypothesis testing.

@@

Popular in-sample tests and model selection criteria, like the Wald
test and the AIC, do not help forecasters in this setting. We show
that these statistics do not select the more accurate model when
choosing between overfit models. For many DGPs the Wald test and the
AIC will choose a larger alternative model over a small benchmark
model with probability approaching one, regardless of which model is
more accurate, while the BIC chooses the benchmark model with
probability approaching one.%
\footnote{Our result holds for a broad class of
  full-sample statistics, but there may be other potential statistics
  that mimic the OOS test and remain valid. Exploring such
  statistics is left for future research.} %
This result holds even though modifications of the $F$-test are valid
under this asymptotic theory, as shown by \citet{BoB:95},
\citet{AkA:00}, \citet{AkP:04}, \citet{Cal:11c}, and \citet{Ana:12},
among others.%
\footnote{Also see \citet{Efr:86,Efr:04} for a discussion
  of naive in-sample loss comparisons.} %
Moreover, under this asymptotic theory, many recent OOS test
statistics, such as those derived by \cite{ClM:01,ClM:05},
\citet{Mcc:07}, and \citet{ClW:06,ClW:07} should have the same
problems as in-sample tests.%
\footnote{Our theoretical results apply
  directly to \citepos{Mcc:07} OOS $t$-test, since it simply
  proposes more liberal critical values for the same test statistic
  that we study.  Since \citet{ClW:06,ClW:07} use a finite length
  estimation window, our asymptotics are incompatible with theirs and
  prevent us from studying their test directly, as well as
  \citepos{GiW:06} and other tests based on \citepos{GiW:06}
  asymptotics.  But \clws\ test can be viewed as a stochastic
  adjustment to the critical values of the usual OOS $t$ test, so
  our conclusions should apply informally as well.  Specifically, we
  show that the DMW test rejects with probability equal to nominal
  size when the estimated benchmark model is expected to be more
  accurate, so more liberal critical values result in overrejection.} %
These tests are also designed to reject the benchmark when the
alternative model is true, and so they may reject too often when the
benchmark is false but more accurate.  Obviously, since the
distribution of these statistics converges to the normal when $P/T \to
0$ (with the number of regressors fixed), these statistics behave like
the DMW test when $P$ is small, but should overreject the benchmark
when $P$ is large. Section~\ref{sec:insample} presents our theoretical
results for full-sample statistics and Section~\ref{sec:mc} presents
Monte Carlo evidence to support these claims.

Finally, this paper introduces a new method of proof for OOS
statistics.  We use a coupling argument (Berbee's Lemma, 1979) to show
that sequences of OOS loss behave like mixingales when the
underlying series are absolutely regular, even if the forecasts depend
on non-convergent estimators.  Moreover, transformations of these
processes also behave like mixingales when recentered, so 
the arguments used to prove asymptotic
results for Near Epoch Dependent (NED) functions of mixing processes
can be used for these OOS processes with only slight modification.

The rest of the paper proceeds as follows.
Section~\ref{sec:assumptions} introduces our notation and assumptions.
Section~\ref{sec:theory} gives the main theoretical results for the
DMW OOS test, shows that standard in-sample tests reject the
benchmark model too often when it is false but more accurate than the
alternative, and shows that standard model selection methods can run
into similar problems. Section~\ref{sec:mc}
presents a Monte Carlo study supporting our theoretical results.
Section~\ref{sec:empirics} applies the OOS statistic to
\citepos{GoW:08} dataset for equity premium prediction, and
Section~\ref{sec:conclusion} concludes.  Proofs and supporting results
are listed in the Appendix.

@@

We assume the following forecasting environment. There are two
competing linear models that give forecasts for the target,
$y_{t+\h}$:
\begin{equation}
y_{t+\h} = x_{1t}'\theta_1 + \e_{1,t+h}, \quad\text{and}\quad
y_{t+\h} = x_{2t}'\theta_2 + \e_{2,t+h};
\end{equation}
$t = 1,\dots,T-h$, $\h$ is the forecast horizon and the
variables $y_t$, $x_{1t}$, and $x_{2t}$ are all known in period $t$.
The coefficients $\theta_1$ and $\theta_2$ minimize the population
Mean Square Error, so
\begin{equation}
  \theta_i = \argmin_\theta \sum_{t=1}^{T-\h} \E (y_{t+\h} - x_{it}'\theta)^2,
\end{equation}
making $\e_{i,t+\h}$ uncorrelated with $x_{i,t}$; $\e_{i,t+\h}$ can exhibit
serial correlation so both of the linear models may be misspecified.
Let $\mathcal{F}_t = \sigma(y_1, x_1, \dots, y_t, x_t)$ be the information
set available in period $t$,
with $x_t$ the vector of all stochastic elements of $x_{1t}$ and $x_{2t}$ after
removing duplicates, and let $\E_t$ and $\var_t$ denote the
conditional mean and variance given $\mathcal{F}_t$.  The first model
uses $K_1$ regressors, and the second uses $K_2$.  Without loss of
generality, assume that $K_1 \leq K_2$. At least one of the models is
overfit, which we represent asymptotically by letting $K_2$ grow with
$T$ quickly enough that $\lim K_2/T$ is positive; $K_1$ may grow with
$T$ as well. Since the models change with $T$, a
stochastic array underlies all of our asymptotic theory, but we
suppress that notation to simplify the presentation.

In the settings we are interested in, a forecaster observes the data
$(y_t,x_t)$ for periods 1 through $T$ and divides the observations
into an estimation sample of the first $R$ observations and a test
sample of the remaining $P$ observations. The forecaster then compares
the models' performance over the test sample, which entails
constructing two sequences of forecasts, (using a fixed-window
estimation strategy)
\begin{equation}
\hat y_{i,t+\h} = x_{it}'\hat{\theta}_{it}, \qquad \text{for } i=1,2;
\ t = R+1,\dots,T-h,
\end{equation}
where
\begin{equation}
  \bh{it} = \Big(\sum_{s=1}^{R-\h} x_{is}x_{is}'\Big)^{-1} \sum_{s=1}^{R-\h}
  x_{is} y_{s+\h}, \qquad \text{for } i=1,2;\ t = R+1,\dots, T - \h.%
\footnote{It may not be clear why we are using the index $t$ in $\bh{it}$,
    since $\bh{it} = \bh{iR}$ almost surely for all $t \leq T-h$.
    But $\bh{it}$ will be defined for $t > T-\h$
    soon and will not equal $\bh{iR}$ for those values of $t$.} %
\end{equation}

The models are then compared by their forecast performance over the test
sample. There are many statistics that have been considered in the
literature, but we focus on perhaps the most natural, the DMW \oost\
test \citep{DiM:95,Wes:96}.%
\footnote{The core insights of our paper apply to other OOS statistics
  as well.} %
This statistic is based on
the difference in the models' loss over the test sample, $\oosA$,
defined as
\begin{equation}
  \oosA \equiv P^{-1} \oosSum{t}{1} D_t
\end{equation}
where
\begin{equation}\label{eq:16}
  D_t = L(y_{t+\h} - x_{1t}'\bh{1t}) - L(y_{t+\h} - x_{2t}'\bh{2t}),
\end{equation}
and $L$ is a known loss function. The \oost\ test is defined as
$\sqrt{P} \oosA / \sh$, where $\sh^2$ is an estimator of the
asymptotic variance of $\oosA$. (Possibly a Heteroskedasticity- and
Autocorrelation-Consistent, or HAC, estimator.)

Statistics like $\oost$ have a long history in empirical economics
because they capture an intuitive idea of modeling: that a good model
should be able to forecast well when given new data and not only when
used to fit the same data that were used to estimate it. \cite{MeR:83}
use such a statistic to study exchange rate models and find that none
of the models that existed at the time of the study outperformed a
random walk. This finding has been remarkably durable and has spawned
an enormous literature; see \cite{Mar:95}, \cite{KiT:03},
\cite{CCP:05}, \cite{EnW:05}, \cite{Ros:05}, and \cite{BWB:10}, among
many others.

Most theoretical research on these statistics, such as \citet{DiM:95}, \citet{Wes:96},
and \cite{Mcc:07}, has focused on using the \oost\ statistic to test
hypotheses about the pseudotrue values $\theta_1$ and $\theta_2$.  In
particular, that research focuses on testing the null hypothesis that
\begin{equation}\label{eq:21}
  \E L(y_{t+\h} - x_{1t}'\theta_{1}) - \E L(y_{t+\h} - x_{2t}'\theta_{2}) = 0.
\end{equation}
But the population quantities in \eqref{eq:21} do not determine which
model is more accurate in practice. The models'
accuracy will also depend on the specific estimates of $\theta_1$ and
$\theta_2$ used to produce the forecasts. In this setting, where the
forecaster will use one of the models to make a number of predictions
in the future---call the number of predictions $Q$---the quantity of
interest becomes
\begin{equation}
  \E_T \oosB = Q^{-1} \sum_{t=T+1}^{T+Q} \E_T D_t,
\end{equation}
where $D_t$ is defined in Equation~\eqref{eq:16} but here uses the
full-sample estimates of the models' parameters,
\begin{equation}
  \bh{it} = \Big(\sum_{s=1}^{T-\h} x_{is}x_{is}'\Big)^{-1} \sum_{s=1}^{T-\h}
  x_{is} y_{s+\h}, \qquad \text{for } i=1,2;\ t = T+1,\dots, T + Q.
\end{equation}
If $\E_T \oosB$ is positive, the second model is expected to forecast
better than the first over the next $Q$ periods, and if $\E_T \oosB$
is negative then the first model is better. We use a conditional
expectation because the coefficient estimates in $\oosB$ are
stochastic but known in period $T$, and their values will determine
the performance of the two models.

Under conventional fixed-$K$ asymptotic theory, $\E_T \oosB$ would
converge in probability to the difference in the expected loss
associated with the pseudotrue models,%
\footnote{This statement is subject to the usual assumptions: some
  form of stationarity, bounded moments, and weak dependence.} %
i.e. the LHS of~\eqref{eq:21}.
\begin{equation}\label{eq:22}
  \E L(y_{t+\h} - x_{1t}'\theta_{1}) - \E L(y_{t+\h} - x_{2t}'\theta_{2}).
\end{equation}
But if $K_2$ increases with $T$ these quantities can have different
limits. For a simple example, assume squared-error loss, let $x_{1,t}
= 1$ for all $t$, and let $(y_{t+\h},x_{2,t})$ be i.i.d. $N(0,
\Sigma)$.  Then the difference between $\E_T \oosB$ and
quantity~\eqref{eq:22} is
\begin{align*}
  \E_T \oosB - & \big(\E L(y_{t+\h} - x_{1t}'\theta_{1}) - \E L(y_{t+\h} - x_{2t}'\theta_{2})\big) \\
  &= \big(\E_T (y_{T+\h+1} - \bh{1,T})^2 - \E_T (y_{T+\h+1} - x_{T+1}'\bh{2,T})^2\big) -
  \big(\E y_{T+\h+1}^2 - \E (y_{T+\h+1} - x_{T+1}'\theta_2)^2 \big)\\
  &= (\bh{2,t} - \theta_2)' \Sigma (\bh{2,t} - \theta_2) + o_p(1).
\end{align*}
This last term has expectation equal to $\var(y_T) \frac{K_2}{T - K_2 - 1}$ and
would converge to zero in probability if $K_2$ were fixed, but does not when $\lim
K_2 / T > 0$. In Section~\ref{sec:oostheory} we show that the \oost\
statistic can estimate $\E_T \oosB$ under our increasing $K$
asymptotics and does not estimate the expected loss associated with
the pseudotrue coefficients.

The conditional expectation $\E_T \oosB$ has been studied heavily in
cross-sectional settings with independent observations. In such a
setting, $\E_T \oosB$ is equal to the difference in the models'
\emph{generalization error}, which has been used widely as a measure
of model accuracy in the machine
learning literature \citep[see][for further discussion]{HTF:08}.
Moreover, with i.i.d. observations, the expectation of $\E_T \oosB$
equals \citepos{Aka:69} Final Prediction Error (FPE). Both
generalization error and FPE are defined by a model's
performance on a new, independent, data set, but, for lack of a better
term, we will call $\E_T \oosB$ the difference in generalization error
for the rest of the paper with hopefully no risk of confusion.

@@

This assumption is a standard condition on the dependence of the
underlying stochastic array. The only novelty is that we use
absolute regularity instead of strong or uniform mixing as our
weak dependence condition; absolute regularity admits a particular
coupling argument, \emph{Berbee's Lemma} \citep[reproduced in this
paper as Lemma A.1 for reference]{Ber:79} that is
unavailable for strong mixing sequences. Absolute regularity
implies uniform mixing, so this assumption is not unduly strong.
For a detailed discussion of these weak dependence conditions,
please see \citet{Dav:94} or \citet{Dou:94}.

@@

  The variance of $y_{t+\h}$ given $\mathcal{F}_t$ is uniformly
  positive and finite and all of the eigenvalues of the covariance
  matrix of $x_t$ are uniformly positive and finite as well.
  Moreover,
  \begin{align*}
    &\eigen_{\max}(X_{iS}'X_{iS}) = O_{L_3}(S), \\
    &\eigen_{\max}((X_{iS}'X_{iS})^{-1}) = O_{L_3}(1/S), \\
    \intertext{and}%
    &\eigen_{\max}\Bigg(\E\Big( \sum_{s,t=U}^{S-h} \e_{i,s+\h} \e_{i,t+\h} x_{is}x_{it}'
    \,\Big|\, X_{iS}'X_{iS} \Big)\Bigg) = O_{L_3}(S-U),
  \end{align*}
  for large enough $T$, where $S = R,\dots,T$, $U=1,\dots,S-\h$,
  $i = 1,2$,
  \[ X_{iS} \equiv [x_{i1} \quad \dots \quad x_{i,S-\h}]' \qquad
  \text{and} \qquad \ep{iS} = (\e_{i,1+\h}, \dots, \e_{i,S})'.\]

@@

The assumption that $y_{t+\h}$ and $x_t$ have positive and finite
variance is straightforward. The conditions on the eigenvalues are
technical and control the behavior of the OLS estimator as the number
of regressors gets large---the third assumption is nonstandard but can
be easily verified under independence, for example. The restrictions
on the pseudotrue coefficients ensure that the regression model
doesn't dominate the variance of $y_{t+\h}$ in the limit.

>>>>

Consider two sequences of length $P$ of prediction errors, the result
of forecasting the same variable with two different estimated models.
Both models are estimated with $R$ observations, collectively called
the {\em estimation window}, and are used to forecast an additional
$P$ observations, called the {\em test sample}.  There are $T$
observations in all, and $R+P=T$.  This paper introduces a new limit
theory for statistics constructed from these prediction errors
designed to approximate the behavior of the statistics when one of the
models is overfit.  In doing so, we provide a theoretical
justification for forecasters to use OOS instead of in-sample
comparisons: the DMW OOS test%
\footnote{In this paper, we will refer to the basic OOS $t$-test
  studied by \citet{DiM:95} and \citet{Wes:96} as the DMW test.} %
allows a forecaster to conduct
inference about the expected future accuracy of his or her models when
one or both is overfit.  We show analytically and through Monte Carlo simulations
that standard full-sample test statistics can not test hypotheses
about this performance.

@@

We focus on linear regression models estimated with a fixed window
for simplicity, but our basic conclusions should be true for other
models and estimation strategies as well.  Under this
asymptotic theory, where the number of regressors $K$ increases with $T$
so that $\lim K/T$ is positive and less than one,
the OLS coefficient estimator is no longer
consistent or asymptotically normal \citep{Hub:73} and has positive
variance in the limit.  We show that, even so, the usual OOS average
is asymptotically normal and can consistently estimate the difference
between the models' generalization error, the expected loss in the
future conditional on the on the data available in period
$T$.%
\footnote{See, for example, \citet{HTF:08} for a discussion of
  generalization error.} %
Under these asymptotics, the generalization
error does not converge to the expected performance of the pseudotrue
models, so existing in-sample and OOS comparisons measure different
quantities and should be expected to give different results for
reasons beyond simple size and power comparisons.  Under our limit theory,
the model that is closer to the true DGP in population can forecast
worse.  In such a situation, a standard in-sample comparison would
correctly reject the null hypothesis that the benchmark is true, and
an OOS comparison would correctly fail to reject the null that the
benchmark is more accurate.%
\footnote{In a pair of papers similar to
  ours, \citet{ClM:12,ClM:12b} study in-sample and OOS tests that
  the larger model has nonzero coefficients that are too close to zero
  to expect the model to forecast more accurately.  Like this paper,
  they argue that the larger model can be true but less accurate.
  However, they focus on an aspect of the DGP that makes this
  phenomenon likely, while we focus on the coefficient estimates that
  produce less accurate forecasts.  Moreover, the implications of our
  asymptotic theories are different and their papers do not provide
  reasons to do OOS comparisons.} %
Also note that, in this situation, a model that performs well
in-sample can perform badly out-of-sample even if there are no
structural breaks or other forms of instability.  Researchers
have argued that a breakdown of in-sample forecasting ability
indicates a structural break (see \citealp{BoH:99}, and
\citealp{StW:03}, among others), but we show that this breakdown
can be caused by overfit as well.

It is important to realize that we are not \emph{advocating} linear
regression for highly overparametrized models. If $K$ is very large
relative to $T$, researchers will often want to use some form of
shrinkage forecast, for example the LASSO \citep{Tib:96} or a factor
model \citep{StW:02,BaN:02}. Our main focus in practice is settings
where $K$ is \emph{moderately} large and it is not clear whether
overfit will dominate the results. In such settings that, it is
important to understand the behavior of different evaluation criteria
when $K$ is large, in case overfit is a concern, and to have reliable
methods for estimating the models' performance, which is the goal of
this paper.

Our theoretical results partially justify OOS comparisons when
researchers want to
choose a model for forecasting.  Although there has been little
emphasis on hypothesis testing in this setting, testing is usually
appropriate: there is usually a familiar benchmark model in place, and
the cost of incorrectly abandoning the benchmark for a less accurate
alternative model is higher than the cost of incorrectly failing to
switch to a more accurate alternative.  We show that the DMW test
lets the forecaster control the probability of the first error, just
as with conventional hypothesis testing.

@@

Popular in-sample tests and model selection criteria, like the Wald
test and the AIC, do not help forecasters in this setting. We show
that these statistics do not select the more accurate model when
choosing between overfit models. For many DGPs the Wald test and the
will choose the larger alternative model over a small benchmark model
with probability much greater than its nominal size, regardless of
which model is more accurate, and the AIC behaves similarly. The BIC,
however, chooses the benchmark model with probability approaching one
even when the \emph{alternative} model is more accurate.%
\footnote{Our result holds for a broad class of full-sample
  statistics, but there may be other potential statistics that mimic
  the OOS test and remain valid. Exploring such statistics is left for
  future research.} %
This result holds even though modifications of the $F$-test are valid
under this asymptotic theory, as shown by \citet{BoB:95},
\citet{AkA:00}, \citet{AkP:04}, \citet{Cal:11c}, and \citet{Ana:12},
among others.%
\footnote{Also see \citet{Efr:86,Efr:04} for a discussion of naive
  in-sample loss comparisons.} %
Moreover, under this asymptotic theory, many recent OOS test
statistics, such as those derived by \cite{ClM:01,ClM:05},
\citet{Mcc:07}, and \citet{ClW:06,ClW:07} should have the same
problems as in-sample tests.%
\footnote{Our theoretical results apply directly to \citepos{Mcc:07}
  OOS $t$-test, since it simply proposes more liberal critical values
  for the same test statistic that we study.  Since
  \citet{ClW:06,ClW:07} use a finite length estimation window, our
  asymptotics are incompatible with theirs and prevent us from
  studying their test directly, as well as \citepos{GiW:06} and other
  tests based on \citepos{GiW:06} asymptotics.  But \clws\ test can be
  viewed as a stochastic adjustment to the critical values of the
  usual \oost\ test, so our conclusions should apply informally as
  well.  Specifically, we show that the DMW test rejects with
  probability equal to nominal size when the estimated benchmark model
  is expected to be more accurate, so more liberal critical values
  result in overrejection.} %
These tests are also designed to reject the benchmark when the
alternative model is true, and so they may reject too often when the
benchmark is misspecified but more accurate.  Obviously, since the
distribution of these statistics converges to the normal when $P/T \to
0$ (with the number of regressors fixed), these statistics behave like
the DMW test when $P$ is small, but should overreject the benchmark
when $P$ is large. Section~\ref{sec:insample} presents our theoretical
results for full-sample statistics and Section~\ref{sec:mc} presents
Monte Carlo evidence to support these claims.

Finally, this paper introduces a new method of proof for OOS
statistics.  We use a coupling argument (Berbee's Lemma, 1979) to show
that sequences of OOS loss behave like mixingales when the
underlying series are absolutely regular, even if the forecasts depend
on non-convergent estimators.  Moreover, transformations of these
processes also behave like mixingales when recentered, so 
many arguments used to prove asymptotic
results for Near Epoch Dependent (NED) functions of mixing processes
can be used for these OOS processes with only slight modification.

The rest of the paper proceeds as follows.
Section~\ref{sec:assumptions} introduces our notation and assumptions.
Section~\ref{sec:theory} gives the main theoretical results for the
DMW OOS test, shows that standard in-sample tests reject the
benchmark model too often when it is misspecified but more accurate than the
alternative, and shows that standard model selection methods can run
into similar problems. Section~\ref{sec:mc}
presents a Monte Carlo study supporting our theoretical results.
Section~\ref{sec:empirics} applies the OOS statistic to
\citepos{GoW:08} dataset for equity premium prediction, and
Section~\ref{sec:conclusion} concludes.  Proofs and supporting results
are listed in the Appendix.

@@

We assume the following forecasting environment. There are two
competing linear models that give forecasts for the target,
$y_{t+\h}$:
\begin{equation*}
y_{t+\h} = x_{1t}'\theta_1 + \e_{1,t+h}, \quad\text{and}\quad
y_{t+\h} = x_{2t}'\theta_2 + \e_{2,t+h};
\end{equation*}
$t = 1,\dots,T-h$, $\h$ is the forecast horizon and the
variables $y_t$, $x_{1t}$, and $x_{2t}$ are all known in period $t$.
The coefficients $\theta_1$ and $\theta_2$ minimize the population
Mean Square Error, so
\begin{equation*}
  \theta_i = \argmin_\theta \sum_{t=1}^{T-\h} \E (y_{t+\h} - x_{it}'\theta)^2,
\end{equation*}
making $\e_{i,t+\h}$ uncorrelated with $x_{i,t}$; $\e_{i,t+\h}$ can exhibit
serial correlation so both of the linear models may be misspecified.
Let $\mathcal{F}_t = \sigma(y_1, x_1, \dots, y_t, x_t)$ be the information
set available in period $t$,
with $x_t$ the vector of all stochastic elements of $x_{1t}$ and $x_{2t}$ after
removing duplicates, and let $\E_t$ and $\var_t$ denote the
conditional mean and variance given $\mathcal{F}_t$.  The first model
uses $K_1$ regressors, and the second uses $K_2$.  Without loss of
generality, assume that $K_1 \leq K_2$. At least one of the models is
overfit, which we represent asymptotically by letting $K_2$ grow with
$T$ quickly enough that $\lim K_2/T$ is positive; $K_1$ may grow with
$T$ as well. Since the models change with $T$, a
stochastic array underlies all of our asymptotic theory, but we
suppress that notation to simplify the presentation.

In the settings we are interested in, a forecaster observes the data
$(y_t,x_t)$ for periods 1 through $T$ and divides these observations
into an estimation sample of the first $R$ observations and a test
sample of the remaining $P$ observations. The forecaster then compares
the models' performance over the test sample, which entails
constructing two sequences of forecasts with a fixed-window
estimation strategy,
\begin{equation*}
\hat y_{i,t+\h} = x_{it}'\hat{\theta}_{it}, \qquad \text{for } i=1,2;
\ t = R+1,\dots,T-h,
\end{equation*}
where
\begin{equation*}
  \bh{it} = \Big(\sum_{s=1}^{R-\h} x_{is}x_{is}'\Big)^{-1} \sum_{s=1}^{R-\h}
  x_{is} y_{s+\h}, \qquad \text{for } i=1,2;\ t = R+1,\dots, T - \h.%
\footnote{It may not be clear why we are using the index $t$ in $\bh{it}$,
    since $\bh{it} = \bh{iR}$ almost surely for all $t \leq T-h$.
    But $\bh{it}$ will be defined for $t > T-\h$
    soon and will not equal $\bh{iR}$ for those values of $t$.} %
\end{equation*}

The models are then compared by their forecast performance over the test
sample. There are many statistics that have been considered in the
literature, but we focus on perhaps the most natural, the DMW \oost\
test \citep{DiM:95,Wes:96}.%
\footnote{The core insights of our paper apply to other OOS statistics
  as well.} %
This statistic is based on
the difference in the models' loss over the test sample, $\oosA$,
defined as
\begin{equation*}
  \oosA \equiv P^{-1} \oosSum{t}{1} D_t
\end{equation*}
where
\begin{equation*}
  D_t = L(y_{t+\h} - x_{1t}'\bh{1t}) - L(y_{t+\h} - x_{2t}'\bh{2t}),
\end{equation*}
and $L$ is a known loss function. The \oost\ test is defined as
$\sqrt{P} \oosA / \sh$, where $\sh^2$ is an estimator of the
asymptotic variance of $\oosA$. (Possibly a Heteroskedasticity- and
Autocorrelation-Consistent, or HAC, estimator.)

Statistics like $\oost$ have a long history in empirical economics
because they capture an intuitive idea of model fit: that a good model
should be able to forecast well on new data. \cite{MeR:83}
use such a statistic to study exchange rate models and find that none
of the models that existed at the time of their study outperform a
random walk. This finding has been remarkably durable and has spawned
an enormous literature; see \cite{Mar:95}, \cite{KiT:03},
\cite{CCP:05}, \cite{EnW:05}, \cite{Ros:05}, and \cite{BWB:10}, among
many others.

Most theoretical research on these statistics, such as \citet{DiM:95}, \citet{Wes:96},
and \cite{Mcc:07}, has focused on using the \oost\ statistic to test
hypotheses about the pseudotrue values $\theta_1$ and $\theta_2$.  In
particular, that research focuses on testing the null hypothesis that
\begin{equation*}
  \E L(y_{t+\h} - x_{1t}'\theta_{1}) = \E L(y_{t+\h} - x_{2t}'\theta_{2}).
\end{equation*}
But the population quantities in this equation do not determine which
model is more accurate in practice. The models'
accuracy will also depend on the specific estimates of $\theta_1$ and
$\theta_2$ used to produce the forecasts.

When the forecaster will use one of the models to make a number of
predictions (call it $Q$) in the future, the quantity of interest
becomes
\begin{equation*}
  \E_T \oosB = Q^{-1} \sum_{t=T+1}^{T+Q} \E_T D_t,
\end{equation*}
where $D_t$ is defined as before,
\begin{equation*}
  D_t = L(y_{t+\h} - x_{1t}'\bh{1t}) - L(y_{t+\h} - x_{2t}'\bh{2t}),
\end{equation*}
but now uses the full-sample estimates of the models' parameters,
\begin{equation*}
  \bh{it} = \Big(\sum_{s=1}^{T-\h} x_{is}x_{is}'\Big)^{-1} \sum_{s=1}^{T-\h}
  x_{is} y_{s+\h}, \qquad \text{for } i=1,2;\ t = T+1,\dots, T + Q.
\end{equation*}
If $\E_T \oosB$ is positive, the second model is expected to forecast
better than the first over the next $Q$ periods, and if $\E_T \oosB$
is negative then the first model is better. We use a conditional
expectation because the coefficient estimates in $\oosB$ are
stochastic but known in period $T$, and their values will determine
the performance of the two models.

Under conventional fixed-$K$ asymptotic theory, $\E_T \oosB$ would
converge in probability to the difference in the expected loss
associated with the pseudotrue models,%
\footnote{This statement is subject to the usual assumptions: some
  form of stationarity, bounded moments, and weak dependence.} %
\begin{equation}\label{eq:22}
  \E L(y_{t+\h} - x_{1t}'\theta_{1}) - \E L(y_{t+\h} - x_{2t}'\theta_{2}).
\end{equation}
But if $K_2$ increases with $T$ these quantities can have different
limits. For a simple example, assume squared-error loss, let $x_{1,t}
= 1$ for all $t$, and let $(y_{t+\h},x_{2,t})$ be i.i.d. $N(0,
\Sigma)$.  Then the difference between $\E_T \oosB$ and
the in quantity~\eqref{eq:22} is
\begin{align*}
  \E_T \oosB - \big(\E L(y_{t+\h} - x_{1t}'\theta_{1}&) - \E L(y_{t+\h} - x_{2t}'\theta_{2})\big) \\
  &= \big(\E_T (y_{T+\h+1} - \bh{1,T})^2 -
     \E_T (y_{T+\h+1} - x_{T+1}'\bh{2,T})^2\big) \\
     &\quad - \big(\E y_{T+\h+1}^2 - \E (y_{T+\h+1} - x_{T+1}'\theta_2)^2 \big)\\
  &= (\bh{2,t} - \theta_2)' \var(x_{2,t})\, (\bh{2,t} - \theta_2) + o_p(1).
\end{align*}
This last term has expectation equal to $\var(y_T) \frac{K_2}{T - K_2 - 1}$ and
would converge to zero in probability if $K_2$ were fixed, but does not when $\lim
K_2 / T > 0$. In Section~\ref{sec:oostheory} we show that the \oost\
statistic can estimate $\E_T \oosB$ under our increasing $K$
asymptotics and does not estimate the expected loss associated with
the pseudotrue coefficients.

The conditional expectation $\E_T \oosB$ has been studied heavily in
cross-sectional settings with independent observations. In such a
setting, $\E_T \oosB$ is equal to the difference in the models'
\emph{generalization error}, which has been used widely as a measure
of model accuracy in the machine
learning literature \citep[see][for further discussion]{HTF:08}.
Moreover, with i.i.d. observations, the expectation of $\E_T \oosB$
equals \citepos{Aka:69} Final Prediction Error (FPE). Both
generalization error and FPE are defined by a model's
performance on a new, independent, data set, but, for lack of a better
term, we will call $\E_T \oosB$ the ``difference in generalization error''
for the rest of the paper with hopefully no risk of confusion.

@@

This assumption is a standard condition on the dependence of the
underlying stochastic array. The only novelty is that we use
absolute regularity instead of strong or uniform mixing as our
weak dependence condition; absolute regularity admits a particular
coupling argument, \emph{Berbee's Lemma} \citep[reproduced in this
paper as Lemma A.1 for reference]{Ber:79} that is
unavailable for strong mixing sequences. Absolute regularity
implies uniform mixing but is more restrictive than strong mixing,
so this assumption is not unduly strong.
For a detailed discussion of these weak dependence conditions,
please see \citet{Dav:94} or \citet{Dou:94}.

@@

  The variance of $y_{t+\h}$ given $\mathcal{F}_t$ is uniformly
  positive and finite and all of the eigenvalues of the covariance
  matrix of $x_t$ are uniformly positive and finite as well.
  Moreover,
  \begin{equation}
    \eigen_{\max}(X_{iS}'X_{iS}) = O_{L_3}(S),
  \end{equation}
  \begin{equation}
    \eigen_{\max}((X_{iS}'X_{iS})^{-1}) = O_{L_3}(1/S),
  \end{equation}
  \begin{multline}\label{eq:28}
    \eigen_{\max}\Bigg(\E\Big(
    \sum_{s,t=U}^{V-\h} \e_{i,s+\h} \e_{i,t+\h} x_{is}x_{it}'
    \mBig
    x_{i1},\dots,x_{i,U-1};
    \sum_{s=U}^{V-\h} x_{is} x_{is}';
    x_{i,V-\h+1},\dots,x_{i,T-\h}
    \Big)\Bigg) \\
    = O_{L_3}(\max(V-U, K_i)),
  \end{multline}
  and
  \begin{multline}\label{eq:8}
    \tr \E\Big(
    \sum_{s,t=U}^{V-\h} \e_{i,s+\h} \e_{i,t+\h} x_{is}x_{it}'
    \mBig
    x_{i1},\dots,x_{i,U-1};
    \sum_{s=U}^{V-\h} x_{is} x_{is}';
    x_{i,V-\h+1},\dots,x_{i,T-\h}
    \Big)\Bigg) \\
    = O_{L_3}((V-U) \times K_i)
  \end{multline}
  for large enough $T$, where $S = R,\dots,T$, $1 \leq U \leq V - \h
  \leq T - \h$, $i = 1,2$,
  \[ X_{iS} \equiv [x_{i1} \quad \dots \quad x_{i,S-\h}]' \qquad
  \text{and} \qquad \ep{iS} = (\e_{i,1+\h}, \dots, \e_{i,S})'.\]

@@

The assumption on $K_1$ and $K_2$ is crucial to the paper; we assume
that the model complexity grows with $T$ fast enough to break
consistency. This assumption is how we derive an asymptotic
concept of ``overfit.''
